Ms 1

### 2018-01-30 ms1/ms1_display.Rmd
study pop chars

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
```

@@

```{r pop_chars}
pop_chars = read_tsv('tables/study_pop_chars.tsv')

pop_chars = read_tsv('tables/study_pop_chars.tsv') %>%
  spread(key, value) %>%
  (function (df) {
    ns1 = names(df)
    ns2 = str_replace(ns1, '^n_(?!bene|cc)', 'pct_')
    rename(df, !!!setNames(as.list(ns1), ns2))
  }) %>%
  mutate_at(vars(matches('^pct_')), function(x) x / .$n_bene * 100) %>%
  gather('key', 'value', -year)

pop_chars %>%
  mutate(key=factor(key, levels=c('n_bene', 'mean_age', 'sd_age', 'n_cc', 'sd_cc', 'pct_female', 'pct_white', 'pct_dual', 'pct_south', 'pct_midwest', 'pct_west', 'pct_northeast'))) %>%
  mutate(value=case_when(
    .$key == 'n_bene' ~ prettyNum(.$value, big.mark=','),
    .$key %in% c('n_cc', 'sd_cc') ~ format(.$value, digits=2, nsmall=2, scientific=FALSE),
    TRUE ~ format(.$value, digits=2, nsmall=1, scientific=FALSE)
  )) %>%
  spread(year, value) %>%
  kable(caption='Population characteristics')

trend_p = function(x) tidy(t.test(diff(x)))$p.value

pop_chars %>%
  arrange(key, year) %>%
  group_by(key) %>%
  summarize(p_value=trend_p(value)) %>%
  kable(caption='Trends in population characteristics')
```

There are `r scan('tables/n_unique_bene.txt') %>% prettyNum(big.mark=',')` unique beneficiaries.

The top 10 drugs take up `r round(scan('tables/top10_cumf.txt') * 100, 2)`% of all claims.

@@

@@

claims_by_abx %>%
  mutate(panel = factor(case_when(
    .$abx %in% c('pde', 'pde_approp', 'pde_inapprop', 'pde_nodx') ~ 1,
    .$abx %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'tmpsmx', 'levofloxacin') ~ 2,
    .$abx %in% c('amoxclav', 'doxycycline', 'nitrofurantoin', 'clindamycin') ~ 3
  ))) %>%
  filter(!is.na(panel)) %>% # remove metro
  mutate(abx=case_when(
    .$abx == 'pde' ~ 'Overall use',
    .$abx == 'pde_inapprop' ~ 'Inappropriate use',
    .$abx == 'pde_approp' ~ 'Appropriate use',
    .$abx == 'pde_nodx' ~ 'Indeterminate use',
    .$abx == 'amoxclav' ~ 'amox/clav',
    .$abx == 'tmpsmx' ~ 'TMP/SMX',
    TRUE ~ .$abx)
  ) %>%
  arrange(year, desc(cpkp)) %>%
  {
  ggplot(., aes(x=year, y=cpkp, group=abx)) +
  geom_blank(data=dummy) +
  geom_line() +
  geom_text_repel(
    data=filter(., year==2013),
    aes(label=abx)
  ) +
  facet_wrap(~panel, scales='free_y', labeller=labeller(panel=panel_labels)) +
  theme_classic() +
  theme(panel.grid.minor.x=element_blank(),
        strip.background = element_blank()) +
  scale_x_continuous(expand=c(0, 0), name='') +
  scale_y_continuous(expand=c(0, 0), name='claims per 1,000 beneficiaries per year')
  }

@@

```{r abx_trends}
est2pct = function(x) (x - 1)*100
format_p = function(p) if_else(p<0.0001, '<.0001', format(round(p, 4), digits=4, nsmall=4))

@@

format_model = function(df) {
  df %>%
    mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
    mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
    mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
    mutate(p.value=format_p(p_value))
}

abx_trends = read_tsv('tables/abx_models.tsv') %>%
  format_model()

@@

abx_trends %>%
  filter(antibiotic != 'metronidazole') %>%
  mutate(antibiotic=case_when(
    .$antibiotic == 'n_pde' ~ 'Overall use',
    .$antibiotic == 'n_pde_inapprop' ~ 'Inappropriate use',
    .$antibiotic == 'n_pde_approp' ~ 'Appropriate use',
    .$antibiotic == 'n_pde_nodx' ~ 'Indeterminate use',
    .$antibiotic == 'amoxclav' ~ 'amox/clav',
    .$antibiotic == 'tmpsmx' ~ 'TMP/SMX',
    TRUE ~ .$antibiotic)
  ) %>%
  # put the abx in order
  arrange(desc(estimate)) %>%
  mutate(antibiotic=fct_inorder(factor(antibiotic))) %>%
  mutate(antibiotic=fct_relevel(antibiotic, c('Overall use', 'Inappropriate use', 'Appropriate use', 'Indeterminate use'))) %>%
  mutate(antibiotic=fct_rev(antibiotic)) %>%
  # plot
  ggplot(aes(x=antibiotic, y=estimate, ymin=lower_ci, ymax=upper_ci)) +
  geom_bar(stat='identity', fill='white', color='black') +
  geom_errorbar(width=0.5) +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in claims per beneficiary per year from 2011 to 2015 (%)') +
  scale_y_continuous(limits=c(-22, 33), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=12),
        axis.text = element_text(size=12))

@@

@@

combined_table = rxdx_table %>%
  mutate(cpkp=round(cpkp, 1)) %>%
  filter(year %in% c(2011, 2015)) %>%
  select(antibiotic, dx_cat, year, cpkp) %>%
  spread(year, cpkp) %>%
  right_join(rxdx_models, by=c('antibiotic', 'dx_cat')) %>%
  format_model() %>%
  select(dx_cat, antibiotic, `2011`, `2015`, change, p.value) %>%
  arrange(dx_cat, antibiotic)
```

@@

subpop_models = read_tsv('tables/subpop_models.tsv') %>%
  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
  mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
  mutate(p.value=format_p(p_value)) %>%
  select(subpop, change, p.value)

@@

```{r ssti}
rxdx_kable(c('cephalexin', 'tmpsmx'),
           c('ssti', 'ssti_t3'))
```

>>>>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)

est2pct = function(x) (x - 1)*100
```

@@

```{r pop_chars}
pop_chars = read_tsv('tables/study_pop_chars.tsv') %>%
  spread(key, value) %>%
  (function (df) {
    ns1 = names(df)
    ns2 = str_replace(ns1, '^n_(?!bene|cc)', 'pct_')
    rename(df, !!!setNames(as.list(ns1), ns2))
  }) %>%
  mutate_at(vars(matches('^pct_')), function(x) x / .$n_bene * 100) %>%
  gather('key', 'value', -year)

pop_chars %>%
  mutate(key=factor(key, levels=c('n_bene', 'mean_age', 'sd_age', 'n_cc', 'sd_cc', 'pct_female', 'pct_white', 'pct_dual', 'pct_south', 'pct_midwest', 'pct_west', 'pct_northeast'))) %>%
  mutate(value=case_when(
    .$key == 'n_bene' ~ prettyNum(.$value, big.mark=','),
    .$key %in% c('n_cc', 'sd_cc') ~ format(.$value, digits=2, nsmall=2, scientific=FALSE),
    TRUE ~ format(.$value, digits=2, nsmall=1, scientific=FALSE)
  )) %>%
  spread(year, value) %>%
  kable(caption='Population characteristics')
```

```{r popchar_models}
pop_char_models = read_tsv('tables/pop_char_models.tsv')

pop_char_models %>%
  mutate_at(vars(estimate, lower_ci, upper_ci), function(x) prettyNum(est2pct(x), digits=3, width=0, format='fg')) %>%
  kable(caption='Trends in study population characteristics')
```

@@

The change in overall use is so small that I'll give it another decimal place:

```{r}
claims_by_abx %>%
  filter(abx=='pde', year %in% c(2011, 2015)) %>%
  select(year, cpkp) %>%
  mutate(cpkp=round(cpkp, 1)) %>%
  kable(caption='Overall use')
```

@@

claims_by_abx %>%
  mutate(panel = factor(case_when(
    .$abx %in% c('pde', 'pde_approp', 'pde_inapprop', 'pde_nodx') ~ 1,
    .$abx %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'tmpsmx', 'levofloxacin') ~ 2,
    .$abx %in% c('amoxclav', 'doxycycline', 'nitrofurantoin', 'clindamycin') ~ 3
  ))) %>%
  filter(!is.na(panel)) %>% # remove metro
  mutate(abx=case_when(
    .$abx == 'pde' ~ 'Overall use',
    .$abx == 'pde_inapprop' ~ 'Inappropriate use',
    .$abx == 'pde_approp' ~ 'Appropriate use',
    .$abx == 'pde_nodx' ~ 'Indeterminate use',
    .$abx == 'amoxclav' ~ 'amox/clav',
    .$abx == 'tmpsmx' ~ 'TMP/SMX',
    TRUE ~ .$abx)
  ) %>%
  arrange(year, desc(cpkp)) %>%
  {
  ggplot(., aes(x=year, y=cpkp, group=abx)) +
  geom_blank(data=dummy) +
  geom_line() +
  geom_text_repel(
    data=filter(., year==2013),
    aes(label=abx)
  ) +
  facet_wrap(~panel, scales='free_y', labeller=labeller(panel=panel_labels)) +
  theme_classic() +
  theme(panel.grid.minor.x=element_blank(),
        strip.background = element_blank(),
        plot.margin=unit(c(1.0, 1.0, 0.5, 0.5), 'lines')) +
  scale_x_continuous(expand=c(0, 0), name='') +
  scale_y_continuous(expand=c(0, 0), name='claims per 1,000 beneficiaries per year')
  }

@@

```{r abx_trends}
format_p = function(p) if_else(p<0.0001, '<.0001', format(round(p, 4), digits=4, nsmall=4))

@@

format_model = function(df) {
  df %>%
    mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
    mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
    #mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
    mutate(change=sprintf('%.1f (%.1f -- %.1f)', estimate, lower_ci, upper_ci)) %>%
    mutate(p.value=format_p(p_value))
}

abx_trends_raw = read_tsv('tables/abx_models.tsv')

abx_trends = abx_trends_raw %>%
  format_model()

@@

abx_trends %>%
  filter(antibiotic != 'metronidazole') %>%
  mutate(antibiotic=case_when(
    .$antibiotic == 'n_pde' ~ 'Overall use',
    .$antibiotic == 'n_pde_inapprop' ~ 'Inappropriate use',
    .$antibiotic == 'n_pde_approp' ~ 'Appropriate use',
    .$antibiotic == 'n_pde_nodx' ~ 'Indeterminate use',
    .$antibiotic == 'amoxclav' ~ 'amox/clav',
    .$antibiotic == 'tmpsmx' ~ 'TMP/SMX',
    TRUE ~ .$antibiotic)
  ) %>%
  # put the abx in order
  arrange(desc(estimate)) %>%
  mutate(antibiotic=fct_inorder(factor(antibiotic))) %>%
  mutate(antibiotic=fct_relevel(antibiotic, c('Overall use', 'Inappropriate use', 'Appropriate use', 'Indeterminate use'))) %>%
  mutate(antibiotic=fct_rev(antibiotic)) %>%
  # plot
  ggplot(aes(x=antibiotic, y=estimate, ymin=lower_ci, ymax=upper_ci)) +
  geom_col(fill='white', color='black') +
  geom_errorbar(width=0.5) +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in claims per beneficiary per year from 2011 to 2015 (%)') +
  scale_y_continuous(limits=c(-20, 30), breaks=c(-20, -10, 0, 10, 20, 30), expand=c(0.0, 0.00)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=12),
        axis.text = element_text(size=12))

@@

Again, overall use declines so little that I'll give it one more decimal place:

```{r}
abx_trends_raw %>%
  filter(antibiotic=='n_pde') %>%
  select(-antibiotic) %>%
  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
  #mutate(change=sprintf('%.2f ± %.2f', estimate, ceiling_at(pm, 2))) %>%
  mutate(change=sprintf('%.2f (%.2f -- %.2f)', estimate, lower_ci, upper_ci)) %>%
  mutate(p.value=format_p(p_value)) %>%
  kable
```

@@

combined_table = rxdx_table %>%
  mutate(cpkp=round(cpkp, 1)) %>%
  filter(year %in% c(2011, 2015)) %>%
  select(antibiotic, dx_cat, year, cpkp) %>%
  spread(year, cpkp) %>%
  right_join(rxdx_models, by=c('antibiotic', 'dx_cat')) %>%
  #format_model() %>%
  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
  mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
  mutate(p.value=format_p(p_value)) %>%
  select(dx_cat, antibiotic, `2011`, `2015`, change, p.value) %>%
  arrange(dx_cat, antibiotic)
```

@@

subpop_models = read_tsv('tables/subpop_models.tsv') %>%
  mutate_at(vars(estimate, lower_ci, upper_ci), est2pct) %>%
  mutate(pm=pmax(estimate-lower_ci, upper_ci-estimate)) %>%
  #mutate(change=sprintf('%.1f ± %.1f', estimate, ceiling_at(pm, 1))) %>%
  mutate(change=sprintf('%.2f (%.2f -- %.2f)', estimate, lower_ci, upper_ci)) %>%
  mutate(p.value=format_p(p_value)) %>%
  select(subpop, change, p.value)

@@

```{r ssti}
rxdx_kable(c('cephalexin', 'tmpsmx', 'ciprofloxacin'),
           c('ssti', 'ssti_t3'))
```
